A New Unitary RG-based Auxiliary Model Approach to Strongly-Correlated Systems

### 2022-04-19 notes/notes.tex
completed obtaining matrix element with corrected normalisation factor

We can use the auxiliary model eigenstates \(\psi_\text{aux}\left(\left\{\vec r_k - \vec R_i\right\} \right) \equiv \psi_\text{aux}\left(\vec R_i\right)\) to construct eigenstates of the bulk Hamiltonian. The following combination of the auxiliary model eigenstates satisfies Bloch's theorem~\cite{stoyanova}:
\begin{equation}\begin{aligned}
	\label{bloch-states}
	\Psi_{\vec k}\left(\left\{ \vec r_k \right\} \right) = \sum_{i} e^{i \vec{k}\cdot\vec{R}_i} \psi_\text{aux}\left(\vec R_i\right)
\end{aligned}\end{equation}
To demonstrate that these are indeed eigenstates of the bulk Hamiltonian, we will calculate the matrix elements of the full Hamiltonian between these states:
\begin{equation}\begin{aligned}
	\bra{\Psi_{\vec k^\prime}}H_\text{H-H} \ket{\Psi_{\vec k}} = \sum_{ij}e^{i\left(\vec{k}\cdot\vec{R}_i - \vec{k}^\prime\cdot\vec{R}_j\right)} \bra{\psi_\text{aux}\left(\vec R_i\right)} H_\text{H-H}\ket{\psi_\text{aux,j}} = \sum_{i}e^{i\left(\vec{k} - \vec k^\prime\right)\cdot\vec{R}_i}\sum_j e^{i\vec k^\prime\cdot\left(\vec R_i - \vec{R}_j\right)} \bra{\psi_\text{aux}\left(\vec R_i\right)} H_\text{H-H}\ket{\psi_\text{aux,j}}
\end{aligned}\end{equation}
We define \(\vec L = \vec R_i - \vec R_j\), and sum over \(\vec L\): 
\begin{equation}\begin{aligned}
\bra{\Psi_{\vec k^\prime}}H_\text{H-H} \ket{\Psi_{\vec k}} = \sum_{\vec R_i}e^{i\left(\vec{k} - \vec k^\prime\right)\cdot\vec{R}_i}\sum_{\vec L} e^{i\vec k^\prime\cdot\vec L} \bra{\psi_\text{aux}\left(\vec R_i\right)} H_\text{H-H}\ket{\psi_\text{aux}\left(\vec R_i- \vec L\right)}
\end{aligned}\end{equation}
The sum over \(\vec L\) can be split into two summations, one over the distance \(L\) and another over the directions \(\vec a(L)\). Also note that because of translation symmetry, the matrix element will depend only on the distance \(L\) and not the direction: \(M(L) \equiv \bra{\psi_\text{aux}\left(\vec R_i\right)} H_\text{H-H}\ket{\psi_\text{aux}\left(\vec R_i- \vec L\right)}\).
\begin{equation}\begin{aligned}
	\bra{\Psi_{\vec k^\prime}(\vec r)}H_\text{H-H} \ket{\Psi_{\vec k}(\vec r)} = \sum_{\vec R_i}e^{i\left(\vec{k} - \vec k^\prime\right)\cdot\vec{R}_i}\sum_{L} M(L) \sum_{\vec a(L)}e^{i\vec k^\prime\cdot\vec a(L)} = \delta_{\vec k, \vec k^\prime}\sum_{L} M(L) \xi_{\vec k}\left(L\right)
\end{aligned}\end{equation}
Here, \(\vec a(L)\) sums over all the primitive vectors connecting any single site with all its neighbours at a distance \(L\), and \(\xi_{\vec k} = \sum_{\vec a(L)}e^{i\vec k^\prime\cdot\vec a(L)}\). The Kronecker delta in front ensures that the Hamiltonian \(H_\text{H-H}\) is diagonal in the basis of the constructed eigenstates. The eigenvalue corresponding to \(\ket{\Psi_{\vec k}}\) is \(\sum_{L} M(L) \xi_{\vec k}\left(L\right)\).

That \(\Psi_{\vec k}\) transforms like Bloch functions can be easily verified:
\begin{equation}\begin{aligned}
	\Psi_{\vec k}\left(\left\{\vec r_k + \vec R\right\} \right) = \sum_{\vec R_i} e^{i \vec{k}\cdot\vec{R}_i} \psi_\text{aux}\left(\vec R_i - \vec R\right) = \sum_{\vec R_j} e^{i \vec{k}\cdot\left(\vec{R}_j + \vec R\right)} \psi_\text{aux}\left(\vec R_j\right) = e^{i \vec{k}\cdot\vec{R}} \Psi_{\vec k}\left(\vec r\right)
\end{aligned}\end{equation}
In the last equation, we transformed \(\vec R_i \to \vec R_j = \vec R_i - \vec R\). The wavefunction can be even brought into the familiar Bloch function form:
\begin{equation}\begin{aligned}
	\Psi_{\vec k}\left(\left\{\vec r_k\right\}\right) = e^{i \vec{k}\cdot\frac{1}{N}\sum_k \vec{r}_k}\sum_{\vec R_i} e^{-i \vec{k}\cdot\left(\frac{1}{N}\sum_k\vec r_k - \vec{R}_i\right)} \psi_\text{aux}\left(\left\{\vec r_k - \vec R_i\right\}\right) = e^{i \vec{k}\cdot\vec{r}_\text{COM}} \eta_{\vec k}\left(\left\{\vec r_k\right\}\right)
\end{aligned}\end{equation}
where \(\vec r_\text{COM} = \frac{1}{N}\sum_k \vec r_k\) is the center-of-mass coordinate and \(\eta_{\vec k}\left(\left\{\vec r_k\right\}\right) = \sum_{\vec R_i} e^{-i \vec{k}\cdot\left(\frac{1}{N}\sum_k\vec r_k - \vec{R}_i\right)} \psi_\text{aux}\left(\left\{\vec r_k - \vec R_i\right\}\right)\) is the translation symmetric function. This form of the eigenstate allows the interpretation that tuning the Bloch momentum \(\vec k\) corresponds to a translation of the center of mass of the system (or in this case, of the auxiliary models that comprise the system).

@@

We define \(\ket{\Phi_0(\vec k)}\) as the ground state of the bulk model, and it has an expression in terms of the auxiliary model eigenstates: \(\ket{\Phi_0(\vec k)} = \sum_i e^{i \vec{k}\cdot\vec{R}_i} \psi_\text{aux}\left( \vec R_i \right) \). To obtain Greens functions, we start by calculating real-space matrix elements of the bulk and auxiliary Hamiltonians. The states that we will consider are local excitations on top of the ground state: \(\ket{i} \equiv c^\dagger_{i\sigma}\ket{\Phi_0\left( \vec k \right) }\). \(i\) represents a lattice site. Sandwiching eq.~\ref{siam_to_hubb} between \(\ket{i}\) and \(\bra{i}\), we get

@@

	\braket{i | H_\text{H-H} | i} = \sum_{\vec R_j} \braket{\Phi_0\left( \vec k \right) | c_{i\sigma} \mathcal{H}_\text{aux}\left( \vec R_j \right) c^\dagger_{i\sigma} | \Phi_0\left( \vec k \right) } = \sum_{\vec R_j, \vec R_m, \vec R_n} e^{i\vec{k}\cdot\left(\vec{R}_m - \vec R_n\right)} \braket{\psi_\text{aux}\left( \vec R_n \right) | c_{i\sigma} \mathcal{H}_\text{aux}\left( \vec R_j \right) c^\dagger_{i\sigma}  | \psi_\text{aux}\left( \vec R_m \right)}

@@

	\ket{\psi_\text{aux}\left( \vec R_m \right)} = \ket{\psi_\text{aux}\left( \left\{\vec r_k - \vec R_m\right\} \right)} = \ket{\psi_\text{aux}\left( \left\{\vec r_k - \vec R_0 + \left(\vec R_0 - \vec R_m\right)\right\} \right)} = T\left( \vec R_0 - \vec R_m \right) \ket{\psi_\text{aux}\left( \left\{\vec r_k - \vec R_0\right\} \right)} \\
	= T\left(\vec R_0 - \vec R_m\right) \ket{\psi_\text{aux}\left(\vec R_0\right)}
	\ket{\psi_\text{aux}\left( \vec R_n \right)} = T\left(\vec R_0 - \vec R_n\right) \ket{\psi_\text{aux}\left(\vec R_0\right)}

@@

	&= \sum_{\vec R_j, \vec R_m, \vec R_n} e^{i\vec{k}\cdot\left(\vec{R}_m - \vec R_n\right)} \braket{\psi_\text{aux}\left( \vec R_0 \right) | T^\dagger\left( \vec R_0 - \vec R_n \right) c_{i\sigma} T\left( \vec R_0 - \vec R_j \right)\mathcal{H}_\text{aux}\left( \vec R_0 \right)T^\dagger\left( \vec R_0 - \vec R_j \right) c^\dagger_{i\sigma}  T\left( \vec R_0 - \vec R_m \right)| \psi_\text{aux}\left( \vec R_0 \right)}

@@

												    &=T^\dagger\left(\vec R_j - \vec R_n\right) c_{\sigma}\left( \vec R_i - \vec R_j + \vec R_0\right)

@@

	T^\dagger\left( \vec R_0 - \vec R_j \right) c^\dagger_{i\sigma} T\left( \vec R_0 - \vec R_m \right) = \left[T^\dagger\left( \vec R_0 - \vec R_m \right) c_{i\sigma} T\left( \vec R_0 - \vec R_j \right)\right]^\dagger = c^\dagger_{\sigma}\left( \vec R_i - \vec R_j + \vec R_0\right) T\left(\vec R_j - \vec R_m\right)
Substituting these forms gives

@@

	&= \sum_{\vec R_j, \vec R_m, \vec R_n} e^{i\vec{k}\cdot\left(\vec{R}_m - \vec R_n\right)} \braket{\psi_\text{aux}\left( \vec R_0 \right) | T^\dagger\left(\vec R_j - \vec R_n\right) c_{\sigma}\left( \vec R_i - \vec R_j + \vec R_0\right)\mathcal{H}_\text{aux}\left( \vec R_0 \right)c^\dagger_{\sigma}\left( \vec R_i - \vec R_j + \vec R_0\right) T\left(\vec R_j - \vec R_m\right)| \psi_\text{aux}\left( \vec R_0 \right)}

@@

@@

	=& \sum_{\vec R_j, \vec R_m, \vec R_n} e^{i\vec{k}\cdot\left(\vec{R}_m - \vec R_n\right)} \bra{\psi_\text{aux}\left( \vec R_0 \right)} \sum_{\vec k_2}\ket{\vec k_2}\bra{\vec k_2}T^\dagger\left(\vec R_j - \vec R_n\right) c_{\sigma}\left( \vec R_i - \vec R_j + \vec R_0\right)\mathcal{H}_\text{aux}\left( \vec R_0 \right)c^\dagger_{\sigma}\left( \vec R_i - \vec R_j + \vec R_0\right)\\
	 &\sum_{\vec k_1}T\left(\vec R_j - \vec R_m\right)\ket{\vec k_1}\bra{\vec k_1} \ket{\psi_\text{aux}\left( \vec R_0 \right)}
From the definition of the translation operators shown above, we get \(T\left(\vec R_j - \vec R_m\right)\ket{\vec k_1} = e^{i \vec{k}_1\cdot\left(\vec{R}_j - \vec R_m\right)}\ket{\vec k_1}\) and \(\bra{\vec k_2}T^\dagger\left(\vec R_j - \vec R_n\right) = \bra{\vec k_2}e^{-i \vec{k}_2\cdot\left(\vec{R}_j - \vec R_n\right)}\). Substituting this gives
	=& \sum_{\vec R_j, \vec R_m, \vec R_n}\sum_{\vec k_1,\vec k_2} e^{i\vec{R}_j\cdot\left(\vec{k}_1 - \vec k_2\right)} e^{i\vec{R}_n\cdot\left(\vec{k}_2 - \vec k\right)} e^{i\vec{R}_m\cdot\left(\vec{k} - \vec k_1\right)} \bra{\psi_\text{aux}\left( \vec R_0 \right)} \ket{\vec k_2}\bra{\vec k_2} c_{\sigma}\left( \vec R_i - \vec R_j + \vec R_0\right)\mathcal{H}_\text{aux}\left( \vec R_0 \right)c^\dagger_{\sigma}\left( \vec R_i - \vec R_j + \vec R_0\right)\\
	 &\ket{\vec k_1}\bra{\vec k_1} \ket{\psi_\text{aux}\left( \vec R_0 \right)}\\
	=& \sum_{\vec R_j}\sum_{\vec k_1,\vec k_2} e^{i\vec{R}_j\cdot\left(\vec{k}_1 - \vec k_2\right)} \delta_{\vec{k},\vec k_2} \delta_{\vec{k},\vec k_1} \bra{\psi_\text{aux}\left( \vec R_0 \right)} \ket{\vec k_2}\bra{\vec k_2} c_{\sigma}\left( \vec R_i - \vec R_j + \vec R_0\right)\mathcal{H}_\text{aux}\left( \vec R_0 \right)c^\dagger_{\sigma}\left( \vec R_i - \vec R_j + \vec R_0\right)\ket{\vec k_1}\bra{\vec k_1} \ket{\psi_\text{aux}\left( \vec R_0 \right)}\\
	=& \sum_{\vec R_j} \braket{\psi_\text{aux}\left( \vec R_0 \right) | \vec k}\bra{\vec k} c_{\sigma}\left( \vec R_i - \vec R_j + \vec R_0\right)\mathcal{H}_\text{aux}\left( \vec R_0 \right)c^\dagger_{\sigma}\left( \vec R_i - \vec R_j + \vec R_0\right)\ket{\vec k}\braket{\vec k | \psi_\text{aux}\left( \vec R_0 \right)}\\

@@

	\braket{i | H_\text{H-H} | i} = |\braket{\psi_\text{aux}\left( \vec R_0 \right) | \vec k}|^2 \bra{\vec k} \sum_{\vec R_j} c_{\sigma}\left(\vec R_j\right)\mathcal{H}_\text{aux}\left( \vec R_0 \right)c^\dagger_{\sigma}\left(\vec R_j\right)\ket{\vec k}

>>>>

We can use the auxiliary model eigenstates \(\psi^n_\text{aux}\left(\left\{\vec r_k - \vec R_i\right\} \right) \equiv \psi^n_\text{aux}\left(\vec R_i\right)\) to construct eigenstates of the bulk Hamiltonian. \(n=0\) is the ground state of the auxiliary model, and higher values denote the excited states. The following combination of the auxiliary model eigenstates satisfies Bloch's theorem~\cite{stoyanova}:
	\label{bloch-states}
	\Psi^n_{\vec k}\left(\left\{ \vec r_k \right\} \right) = \frac{1}{\sqrt N}\sum_{i} e^{i \vec{k}\cdot\vec{R}_i} \psi^n_\text{aux}\left(\vec R_i\right)
where \(N\) is the total number of lattice sites. The interpretation is that the set of \(n=0\) states form the lowest band in the spectrum of the lattice. Higher values of \(n\) produce the more energetic bands. The ground state is obtained by setting \(\vec k\) and \(n\) to 0: \(\ket{\Psi_{\vec k=0}^{n=0}} \equiv \ket{\Psi_\text{gs}}\)To demonstrate that these are indeed eigenstates of the bulk Hamiltonian, we will calculate the matrix elements of the full Hamiltonian between these states:
	&\bra{\Psi^n_{\vec k^\prime}}H_\text{H-H} \ket{\Psi^n_{\vec k}}\\
	&= \frac{1}{N}\sum_{ij}e^{i\left(\vec{k}\cdot\vec{R}_i - \vec{k}^\prime\cdot\vec{R}_j\right)} \bra{\psi^n_\text{aux}\left(\vec R_i\right)} H_\text{H-H}\ket{\psi^n_\text{aux}\left(\vec R_j\right)} = \frac{1}{N}\sum_{i,j}e^{i\left(\vec{k} - \vec k^\prime\right)\cdot\vec{R}_i} e^{i\vec k^\prime\cdot\left(\vec R_i - \vec{R}_j\right)} \bra{\psi^n_\text{aux}\left(\vec R_i\right)} H_\text{H-H}\ket{\psi^n_\text{aux}\left(\vec R_j\right)}
We define \(\vec L = \vec R_i - \vec R_j\), and sum over \(\vec L\): 
\bra{\Psi^n_{\vec k^\prime}}H_\text{H-H} \ket{\Psi^n_{\vec k}} = \frac{1}{N}\sum_{\vec R_i}e^{i\left(\vec{k} - \vec k^\prime\right)\cdot\vec{R}_i}\sum_{\vec L} e^{i\vec k^\prime\cdot\vec L} \bra{\psi^n_\text{aux}\left(\vec R_i\right)} H_\text{H-H}\ket{\psi^n_\text{aux}\left(\vec R_i- \vec L\right)}
Note that because of translation symmetry, the matrix element will depend only on the distance \(\vec L\) and not the position \(\vec R_i\): \(C(-\vec L) \equiv \bra{\psi^n_\text{aux}\left(\vec R_i\right)} H_\text{H-H}\ket{\psi^n_\text{aux}\left(\vec R_i- \vec L\right)}\).
	\bra{\Psi^n_{\vec k^\prime}(\vec r)}H_\text{H-H} \ket{\Psi^n_{\vec k}(\vec r)} = \frac{1}{N}\sum_{\vec R_i}e^{i\left(\vec{k} - \vec k^\prime\right)\cdot\vec{R}_i}\sum_{\vec L} C(-\vec L) e^{i\vec k^\prime\cdot\vec L} = \frac{1}{N}\delta_{\vec k, \vec k^\prime}\times N\sum_{L} C(\vec L) e^{-i\vec k\cdot\vec L} = \delta_{\vec k, \vec k^\prime} \sum_{L} C(\vec L) e^{-i\vec k\cdot\vec L}
The Kronecker delta in front ensures that the Hamiltonian \(H_\text{H-H}\) is diagonal in the basis of the constructed eigenstates. The eigenvalue corresponding to \(\ket{\Psi^n_{\vec k}}\) is \(\sum_{L} C(\vec L) e^{-i\vec k\cdot\vec L}\). For thr ground state \(\vec k=0,n=0\), this can be simplified into
	\sum_{L} C(\vec L) &= \sum_{\vec L}\bra{\psi^0_\text{aux}\left(\vec R_0\right)} H_\text{H-H}\ket{\psi^0_\text{aux}\left(\vec R_0 - \vec L\right)} = \sum_{\vec L,\vec R_j}\bra{\psi^0_\text{aux}\left(\vec R_0\right)} H_\text{aux}\left(\vec R_j\right) T\left( \vec L \right) \ket{\psi^0_\text{aux}\left(\vec R_0\right)}\\
			   &=\sum_{\vec L,\vec R_j}\bra{\psi^0_\text{aux}\left(\vec R_0\right)} T(\vec R_0 - \vec R_j) H_\text{aux}\left(\vec R_0\right)T^\dagger(\vec R_0 - \vec R_j) T\left( \vec L \right) \ket{\psi^0_\text{aux}\left(\vec R_0\right)}\\
			   &=\sum_{\vec L,\vec R_j}\bra{\psi^0_\text{aux}\left(\vec R_0\right)} T(\vec R_0 - \vec R_j) H_\text{aux}\left(\vec R_0\right) T\left(\vec L + \vec R_j - \vec R_0\right) \ket{\psi^0_\text{aux}\left(\vec R_0\right)}\\
To resolve the remaining translation operators, we will now insert the eigenstates \(\ket{\Psi_{\vec k}}\) of the full model: \(1 = \sum_{\vec k,n}\ket{\Psi^n_{\vec k}}\bra{\Psi^n_{\vec k}}\) in between the operators. These eigenkets can in general be any state with a definite momentum, the simplest case being the eigenstates of a tight-binding Hamiltonian obtained by setting \(U=J=0\) in the bulk model. We have chosen the eigenstates of the full model for our purpose. We already know, from Bloch's theorem, that \(T(\vec R)\ket{\Psi_{\vec k}} = e^{i \vec{k}\cdot\vec{R}}\ket{\Psi_{\vec k}}\). 
	\sum_{L} C(\vec L) &= \sum_{n,\vec L,\vec R_j, \vec k_1, \vec k_2}\bra{\psi^0_\text{aux}\left(\vec R_0\right)} \ket{\Psi^n_{\vec k_2}}\bra{\Psi^n_{\vec k_2}} T^\dagger(-\vec R_0 + \vec R_j) H_\text{aux}\left(\vec R_0\right) T\left(\vec L + \vec R_j - \vec R_0\right) \ket{\Psi^n_{\vec k_1}}\bra{\Psi^n_{\vec k_1}}\ket{\psi^0_\text{aux}\left(\vec R_0\right)}\\
			   &= \sum_{n,\vec L,\vec R_j, \vec k_1, \vec k_2}\bra{\psi^0_\text{aux}\left(\vec R_0\right)} \ket{\Psi^n_{\vec k_2}}\bra{\Psi^n_{\vec k_2}} e^{-i \vec k_2 \cdot \left(-\vec R_0 + \vec R_j\right)} H_\text{aux}\left(\vec R_0\right) e^{i \vec k_1 \cdot \left(\vec L + \vec R_j - \vec R_0\right)} \ket{\Psi^n_{\vec k_1}}\bra{\Psi^n_{\vec k_1}}\ket{\psi^0_\text{aux}\left(\vec R_0\right)}\\
			   &= \sum_{n,\vec L,\vec k_1, \vec k_2}e^{i \vec{\Psi^n_{\vec k_1}}\cdot\vec{L}}\bra{\psi^0_\text{aux}\left(\vec R_0\right)} \ket{\Psi^n_{\vec k_2}}\bra{\Psi^n_{\vec k_2}} N\delta_{\vec k_1,\vec k_2} H_\text{aux}\left(\vec R_0\right) \ket{\Psi^n_{\vec k_1}}\bra{k_1}\ket{\psi^0_\text{aux}\left(\vec R_0\right)}\\
			   &= N\sum_{n,\vec L, \vec k}e^{i \vec{k}\cdot\vec{L}} |\braket{\psi^0_\text{aux}\left(\vec R_0\right) | \Psi_{\vec k}^n}|^2 \bra{\Psi_{\vec k}^n} H_\text{aux}\left(\vec R_0\right) \ket{\Psi_{\vec k}^n}\\
			   &= N^2 |\braket{\psi^0_\text{aux}\left(\vec R_0\right) | \Psi_\text{gs}}|^2 \bra{\Psi_\text{gs}} H_\text{aux}\left(\vec R_0\right) \ket{\Psi_\text{gs}}\\
\(\ket{\Psi_\text{gs}}\) is the ground state of the bulk model, and can be expressed as a sum over the auxiliary models at varying distances from \(\vec R_0\). This form allows a systematic method of improving the energy eigenvalue estimate, because the interacting in the impurity model is extremely localised, so the overlap between two auxiliary models decreases rapidly with distance. The majority of the contribution will come from the auxiliary model at \(\vec R_0\), and improvements are then made by considering auxiliary models at further distances.

It is easy to verify that \(\Psi^n_{\vec k}\) transforms like Bloch functions:
\begin{equation}\begin{aligned}
	\Psi^n_{\vec k}\left(\left\{\vec r_k + \vec R\right\} \right) = \frac{1}{\sqrt N}\sum_{\vec R_i} e^{i \vec{k}\cdot\vec{R}_i} \psi^n_\text{aux}\left(\vec R_i - \vec R\right) = \frac{1}{\sqrt N}\sum_{\vec R_j} e^{i \vec{k}\cdot\left(\vec{R}_j + \vec R\right)} \psi^n_\text{aux}\left(\vec R_j\right) = e^{i \vec{k}\cdot\vec{R}} \Psi^n_{\vec k}\left(\vec r\right)
\end{aligned}\end{equation}
In the last equation, we transformed \(\vec R_i \to \vec R_j = \vec R_i - \vec R\). The wavefunction can be even brought into the familiar Bloch function form:
\begin{equation}\begin{aligned}
	\Psi^n_{\vec k}\left(\left\{\vec r_k\right\}\right) = \frac{1}{\sqrt N}e^{i \vec{k}\cdot\frac{1}{N}\sum_k \vec{r}_k}\sum_{\vec R_i} e^{-i \vec{k}\cdot\left(\frac{1}{N}\sum_k\vec r_k - \vec{R}_i\right)} \psi^n_\text{aux}\left(\left\{\vec r_k - \vec R_i\right\}\right) = e^{i \vec{k}\cdot\vec{r}_\text{COM}} \eta_{\vec k}\left(\left\{\vec r_k\right\}\right)
\end{aligned}\end{equation}
where \(\vec r_\text{COM} = \frac{1}{N}\sum_k \vec r_k\) is the center-of-mass coordinate and \(\eta_{\vec k}\left(\left\{\vec r_k\right\}\right) = \frac{1}{\sqrt N}\sum_{\vec R_i} e^{-i \vec{k}\cdot\left(\frac{1}{N}\sum_k\vec r_k - \vec{R}_i\right)} \psi^n_\text{aux}\left(\left\{\vec r_k - \vec R_i\right\}\right)\) is the translation symmetric function. This form of the eigenstate allows the interpretation that tuning the Bloch momentum \(\vec k\) corresponds to a translation of the center of mass of the system (or in this case, of the auxiliary models that comprise the system).

@@

In the previous part, we obtained, among other states, the ground state \(\ket{\Psi_\text{gs}} = \ket{\Psi^{n=0}_{\vec k=0}} = \frac{1}{\sqrt N}\sum_{\vec R}\ket{\psi^0_\text{aux}\left( \vec R \right)}\) of the bulk Hamiltonian.
To obtain Greens functions, we start by calculating real-space matrix elements of the bulk and auxiliary Hamiltonians. The states that we will consider are local excitations on top of the ground state: \(\ket{i} \equiv c^\dagger_{i\sigma}\ket{\Psi_\text{gs} }\). \(i\) represents a lattice site. Sandwiching eq.~\ref{siam_to_hubb} between \(\ket{i}\) and \(\bra{i}\), we get

@@

	\braket{i | H_\text{H-H} | i} = \sum_{\vec R_j} \braket{\Psi_\text{gs} | c_{i\sigma} \mathcal{H}_\text{aux}\left( \vec R_j \right) c^\dagger_{i\sigma} | \Psi_\text{gs} } = \frac{1}{N}\sum_{\vec R_j, \vec R_m, \vec R_n} e^{i\vec{k}\cdot\left(\vec{R}_m - \vec R_n\right)} \braket{\psi^0_\text{aux}\left( \vec R_n \right) | c_{i\sigma} \mathcal{H}_\text{aux}\left( \vec R_j \right) c^\dagger_{i\sigma}  | \psi^0_\text{aux}\left( \vec R_m \right)}

@@

	\ket{\psi^0_\text{aux}\left( \vec R_m \right)} = \ket{\psi^0_\text{aux}\left( \left\{\vec r_k - \vec R_m\right\} \right)} = \ket{\psi^0_\text{aux}\left( \left\{\vec r_k - \vec R_0 + \left(\vec R_0 - \vec R_m\right)\right\} \right)} = T\left( \vec R_0 - \vec R_m \right) \ket{\psi^0_\text{aux}\left( \left\{\vec r_k - \vec R_0\right\} \right)} \\
	= T\left(\vec R_0 - \vec R_m\right) \ket{\psi^0_\text{aux}\left(\vec R_0\right)}
	\ket{\psi^0_\text{aux}\left( \vec R_n \right)} = T\left(\vec R_0 - \vec R_n\right) \ket{\psi^0_\text{aux}\left(\vec R_0\right)}

@@

	&= \frac{1}{N}\sum_{\vec R_j, \vec R_m, \vec R_n} e^{i\vec{k}\cdot\left(\vec{R}_m - \vec R_n\right)} \braket{\psi^0_\text{aux}\left( \vec R_0 \right) | T^\dagger\left( \vec R_0 - \vec R_n \right) c_{i\sigma} T\left( \vec R_0 - \vec R_j \right)\mathcal{H}_\text{aux}\left( \vec R_0 \right)T^\dagger\left( \vec R_0 - \vec R_j \right) c^\dagger_{i\sigma}  T\left( \vec R_0 - \vec R_m \right)| \psi^0_\text{aux}\left( \vec R_0 \right)}

@@

												    &=T^\dagger\left(\vec R_j - \vec R_n\right) c_{\vec R_i - \vec R_j + \vec R_0,\sigma}

@@

	T^\dagger\left( \vec R_0 - \vec R_j \right) c^\dagger_{i\sigma} T\left( \vec R_0 - \vec R_m \right) = \left[T^\dagger\left( \vec R_0 - \vec R_m \right) c_{i\sigma} T\left( \vec R_0 - \vec R_j \right)\right]^\dagger = c^\dagger_{\vec R_i - \vec R_j + \vec R_0,\sigma} T\left(\vec R_j - \vec R_m\right)
The operator \(c^\dagger_{\vec R_i - \vec R_j + \vec R_0,\sigma}\) creates an electron of spin \(\sigma\) at the position \(\vec R_i - \vec R_j + \vec R_0\). It will be represented as \(c^\dagger_{\vec R_i - \vec R_j + \vec R_0,\sigma}\). Substituting these forms gives

@@

	&= \frac{1}{N}\sum_{\vec R_j, \vec R_m, \vec R_n} e^{i\vec{k}\cdot\left(\vec{R}_m - \vec R_n\right)} \braket{\psi^0_\text{aux}\left( \vec R_0 \right) | T^\dagger\left(\vec R_j - \vec R_n\right) c_{\vec R_i - \vec R_j + \vec R_0,\sigma}\mathcal{H}_\text{aux}\left( \vec R_0 \right)c^\dagger_{\vec R_i - \vec R_j + \vec R_0,\sigma} T\left(\vec R_j - \vec R_m\right)| \psi^0_\text{aux}\left( \vec R_0 \right)}

@@

	\braket{i | H_\text{H-H} | i} = \frac{1}{N}\sum_{\vec R_j, \vec R_m, \vec R_n} e^{i\vec{k}\cdot\left(\vec{R}_m - \vec R_n\right)} \bra{\psi^0_\text{aux}\left( \vec R_0 \right)} \sum_{\vec k_2}\ket{\vec k_2}\bra{\vec k_2}T^\dagger\left(\vec R_j - \vec R_n\right) c_{\vec R_i - \vec R_j + \vec R_0,\sigma}\mathcal{H}_\text{aux}\left( \vec R_0 \right)c^\dagger_{\vec R_i - \vec R_j + \vec R_0,\sigma}\\
	 \sum_{\vec k_1}T\left(\vec R_j - \vec R_m\right)\ket{\vec k_1}\bra{\vec k_1} \ket{\psi^0_\text{aux}\left( \vec R_0 \right)}
From the definition of the translation operators shown above, we get 
T\left(\vec R_j - \vec R_m\right)\ket{\vec k_1} = e^{i \vec{k}_1\cdot\left(\vec{R}_j - \vec R_m\right)}\ket{\vec k_1}, ~ ~\bra{\vec k_2}T^\dagger\left(\vec R_j - \vec R_n\right) = \bra{\vec k_2}e^{-i \vec{k}_2\cdot\left(\vec{R}_j - \vec R_n\right)}~.
 Substituting this gives

@@

	=& \frac{1}{N} \sum_{\vec R_j, \vec R_m, \vec R_n}\sum_{\vec k_1,\vec k_2} e^{i\vec{R}_j\cdot\left(\vec{k}_1 - \vec k_2\right)} e^{i\vec{R}_n\cdot\left(\vec{k}_2 - \vec k\right)} e^{i\vec{R}_m\cdot\left(\vec{k} - \vec k_1\right)} \bra{\psi^0_\text{aux}\left( \vec R_0 \right)} \ket{\vec k_2}\bra{\vec k_2} c_{\vec R_i - \vec R_j + \vec R_0,\sigma}\mathcal{H}_\text{aux}\left( \vec R_0 \right)c^\dagger_{\vec R_i - \vec R_j + \vec R_0,\sigma}\\
	 &\ket{\vec k_1}\bra{\vec k_1} \ket{\psi^0_\text{aux}\left( \vec R_0 \right)}\\
	=& \frac{1}{N} \sum_{\vec R_j}\sum_{\vec k_1,\vec k_2} e^{i\vec{R}_j\cdot\left(\vec{k}_1 - \vec k_2\right)} N^2 \delta_{\vec{k},\vec k_2} \delta_{\vec{k},\vec k_1} \bra{\psi^0_\text{aux}\left( \vec R_0 \right)} \ket{\vec k_2}\bra{\vec k_2} c_{\vec R_i - \vec R_j + \vec R_0,\sigma}\mathcal{H}_\text{aux}\left( \vec R_0 \right)c^\dagger_{\vec R_i - \vec R_j + \vec R_0,\sigma}\ket{\vec k_1}\bra{\vec k_1} \ket{\psi^0_\text{aux}\left( \vec R_0 \right)}\\
	=& N \sum_{\vec R_j} \braket{\psi^0_\text{aux}\left( \vec R_0 \right) | \vec k}\bra{\vec k} c_{\vec R_i - \vec R_j + \vec R_0,\sigma}\mathcal{H}_\text{aux}\left( \vec R_0 \right)c^\dagger_{\vec R_i - \vec R_j + \vec R_0,\sigma}\ket{\vec k}\braket{\vec k | \psi^0_\text{aux}\left( \vec R_0 \right)}\\

@@

	\braket{i | H_\text{H-H} | i} = N |\braket{\psi^0_\text{aux}\left( \vec R_0 \right) | \vec k}|^2 \bra{\vec k} \sum_{\vec R_j} c_{\sigma}\left(\vec R_j\right)\mathcal{H}_\text{aux}\left( \vec R_0 \right)c^\dagger_{\sigma}\left(\vec R_j\right)\ket{\vec k}
Even though it is computationally easier to consider the state \(\ket{\vec k}\) as the momentum eigenstates of the tight-binding kind, there is a more illuminating interpretation of this expression if we consider them as the eigenstates \(\ket{\Phi\left( \vec k \right) }\) of the full bulk model. With this choice, the expression can be seen as the sum of effects of creating a single-particle excitation \(c^\dagger_\sigma\) on the ground state \(\ket{\Psi_\text{gs} }\), at increasingly larger distances \(\vec R_j\) from the auxiliary model at \(\vec R_0\). The excitation is made to evolve through the auxiliary model \(\mathcal{H}_\text{aux}\left( \vec R_0 \right) \) centred at \(\vec R_0\). The excitation is ultimately annihilated and the state returns to the ground state. However, since we are describing the excitation of the bulk ground state using an auxiliary model Hamiltonian, we also need to multply this with the fidelity between the full ground state and the auxiliary model ground state. That is taken care of by the inner-product multiplying the matrix element. That inner product can be thought of as an overlap between a separable state and an entangled state - the full Hamiltonian state is obtained by entangling the overlapping auxiliary models, while the single auxiliary model groundstate is the separable state in this context. The inner product then becomes a measure of entanglement of the bulk ground state - the more the overlap with the separable state, the lower the entanglement: \(\varepsilon(\vec R_0,\vec k) \equiv 1 - |\braket{\psi^0_\text{aux}\left(  \right) | \Phi_0(\vec k)}|^2\).
	\braket{i | H_\text{H-H} | i} = N \left( 1 - \varepsilon(\vec R_0,\vec k) \right) \bra{\Phi_0(\vec k)} \sum_{\vec R_j} c_{\sigma}\left(\vec R_j\right)\mathcal{H}_\text{aux}\left( \vec R_0 \right)c^\dagger_{\sigma}\left(\vec R_j\right)\ket{\Phi_0(\vec k)}
